### 2018-04-30 tex/tech_occ_manual.tex

\begin{longtable}{ccc} % [!htbp] 
  \caption{Technology Occupations} \label{tab:tab_tech_occ}\\
\hline \\[-1.8ex] 
Type & SOC & Occupation Title \\ 
\hline \\[-1.8ex] 
\endfirsthead
\hline \\[-1.8ex] 
Type & SOC & Occupation Title \\ 
\hline \\[-1.8ex] 
\endhead
\hline \\[-1.8ex] 
\endfoot
\multirow{17}{*}{STEM} & 11-3020 & Computer and Information Systems Managers \\ 
 & 11-9040 & Architectural and Engineering Managers \\ 
 & 11-9120 & Natural Sciences Managers \\ 
 & 15-0000 & Computer and Mathematical Occupations \\ 
 & 17-1020 & Surveyors, Cartographers, and Photogrammetrists  \\ 
 & 17-2000 & Engineers \\ 
 & 17-3000 & Drafters, Engineering Technicians, and Mapping Technicians \\ 
 & 19-1000 & Life Scientists \\ 
 & 19-2000 & Physical Scientists  \\ 
 & 19-3010 & Economists \\ 
 & 19-3022 & Survey Researchers \\ 
 & 19-3030 & Psychologists  \\ 
 & 19-3040 & Sociologists \\ 
 & 19-3050 & Urban and Regional Planners \\ 
 & 19-3090 & Miscellaneous Social Scientists and Related Workers  \\ 
 & 19-4000 & Life, Physical, and Social Science Technicians  \\ 
 & 41-9031 & Sales Engineers \\ 
\hline \\[-1.8ex] 
\multirow{17}{*}{STEM Related} & 11-3020 & Computer and Information Systems Managers \\ 
 & 11-9040 & Architectural and Engineering Managers \\ 
 & 11-9110 & Medical and Health Services Managers  \\ 
 & 11-9120 & Natural Sciences Managers \\ 
 & 15-0000 & Computer and Mathematical Occupations \\ 
 & 17-0000 & Architecture and Engineering Occupations \\ 
 & 19-1000 & Life Scientists \\ 
 & 19-2000 & Physical Scientists  \\ 
 & 19-3010 & Economists \\ 
 & 19-3022 & Survey Researchers \\ 
 & 19-3030 & Psychologists  \\ 
 & 19-3040 & Sociologists \\ 
 & 19-3050 & Urban and Regional Planners \\ 
 & 19-3090 & Miscellaneous Social Scientists and Related Workers  \\ 
 & 19-4000 & Life, Physical, and Social Science Technicians  \\ 
 & 29-0000 & Healthcare Practitioners and Technical Occupations \\ 
 & 41-9031 & Sales Engineers \\ 
 \hline \\[-1.8ex] 
\multirow{12}{*}{High Tech} & 11-3020 & Computer and Information Systems Managers  \\ 
 & 11-9040 & Architectural and Engineering Managers  \\ 
 & 11-9120 & Natural Sciences Managers  \\ 
 & 15-0000 & Computer and Mathematical Occupations \\ 
 & 17-2000 & Engineers  \\ 
 & 17-3000 & Drafters, Engineering Technicians, and Mapping Technicians  \\ 
 & 19-1000 & Life Scientists  \\ 
 & 19-2000 & Physical Scientists \\ 
 & 19-4000 & Life, Physical, and Social Science Technicians \\ 
 & 27-1014 & Multimedia Artists and Animators \\ 
 & 29-2011 & Medical and Clinical Laboratory Technologists  \\ 
 & 29-2033 & Nuclear Medicine Technologists \\ 
 & 29-2034 & Radiologic Technologists \\ 
\hline \\[-1.8ex] 
\multirow{4}{*}{Tech Group} & 15-0000 & Computer and Mathematical Occupations \\ 
 & 17-0000 & Architecture and Engineering Occupations \\ 
 & 19-0000 & Life, Physical, and Social Science Occupations \\ 
 & 29-0000 & Healthcare Practitioners and Technical Occupations \\ 
\end{longtable}