### 2018-04-30 tex/top_bottom_tech_ind_manual.tex

\begin{table}[!htbp] \centering 
  \caption{Top and Bottom Technology Industry Sectors in 2016} 
  \label{tab:tab_top_bottom_tech_ind} 
\begin{tabular}{@{\extracolsep{5pt}} ccc} 
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
NAICS & Title & \thead{Share of \\ Tech Workers} \\ 
\hline \\[-1.8ex] 
541500 & Computer Systems Design and Related Services & 0.642 \\ 
511200 & Software Publishers & 0.591 \\ 
334100 & Computer and Peripheral Equipment Manufacturing & 0.576 \\ 
541700 & Scientific Research and Development Services & 0.570 \\ 
541300 & Architectural, Engineering, and Related Services & 0.544 \\ 
518200 & Data Processing, Hosting, and Related Services & 0.455 \\ 
334200 & Communications Equipment Manufacturing & 0.389 \\ 
334500 & \makecell{Navigational, Measuring, Electromedical, \\ and Control Instruments Manufacturing} & 0.371 \\ 
517400 & Satellite Telecommunications & 0.354 \\ 
334400 & \makecell{Semiconductor and Other Electronic Component Manufacturing} & 0.354 \\ 
\hline \\[-1.8ex] 
316100 & Leather and Hide Tanning and Finishing & 0.000 \\ 
445300 & Beer, Wine, and Liquor Stores  & 0.000 \\ 
453100 & Florists  & 0.000 \\ 
454200 & Vending Machine Operators  & 0.000 \\ 
485200 & Interurban and Rural Bus Transportation & 0.000 \\ 
485300 & Taxi and Limousine Service & 0.000 \\ 
485400 & School and Employee Bus Transportation & 0.000 \\ 
485500 & Charter Bus Industry & 0.000 \\ 
485900 & Other Transit and Ground Passenger Transportation & 0.000 \\ 
487100 & Scenic and Sightseeing Transportation, Land & 0.000 \\ 
487200 & Scenic and Sightseeing Transportation, Water & 0.000 \\ 
487900 & Scenic and Sightseeing Transportation, Other & 0.000 \\ 
492200 & Local Messengers and Local Delivery & 0.000 \\ 
525100 & Insurance and Employee Benefit Funds  & 0.000 \\ 
525900 & Other Investment Pools and Funds & 0.000 \\ 
532300 & General Rental Centers & 0.000 \\ 
721300 & Rooming and Boarding Houses & 0.000 \\ 
722400 & Drinking Places (Alcoholic Beverages) & 0.000 \\ 
811400 & Personal and Household Goods Repair and Maintenance & 0.000 \\ 
\hline \\[-1.8ex] 
\end{tabular} 
\end{table}